---
title: "state-level-DIME"
output: 
  html_document:
    code_folding: hide
---
```{r setup, include=FALSE}
## Defaults for R chunks
knitr::opts_chunk$set(echo = TRUE, ## echo = TRUE means code will show
                      warning=FALSE, ## supress warnings and messages from R
                      message=FALSE,
                      fig.path='Figs/', ## where to save figures
                      fig.height = 3, ## default figure size (can differ in each chunk)
                      fig.width = 3)
## Add any R packages you require. 
## Here are some we will use in 811:
requires <- c("tidyverse", ## tidyverse includes dplyr and ggplot2
              "magrittr",
              "here")
## Install any you don't have
to_install <- c(requires %in% rownames(installed.packages()) == FALSE)
install.packages(c(requires[to_install], "NA"), repos = "https://cloud.r-project.org/" )
## Load all required R packages
library(tidyverse)
library(ggplot2); theme_set(theme_minimal())
library(magrittr)
library(here)
```


## How much ideological diversity is there in U.S. state governments?

Using DIME ideology scores (based on campaign donations).

```{r DIME_data}
## load data from a web address

load("data/BonicaStateLevel.Rdata")

d %<>%
  mutate(year = year -1) %>%
  ## naming things! 
  rename(Governor = gov_cf) %>% 
  rename(Lower = l_cf_median) %>% 
  rename(Upper = u_cf_median) %>%
  mutate(state_ideology = ifelse(dist_1_x_zero == 0 & Governor < 0, "All Liberal", "Mixed")) %>% 
  mutate(state_ideology = ifelse(dist_1_x_zero == 0 & Governor > 0, "All Conservative", state_ideology)) %>%
  mutate(state_ideology = ifelse(is.na(dist_1_x_zero) & Governor > 0, "All Conservative", state_ideology)) %>%
  ## identify min and max DIME scores for each state (each row is a state)
  mutate(min = pmin(Governor, Lower, Upper, na.rm = T), 
         max = pmax(Governor, Lower, Upper, na.rm = T) )


## Legislator data only
# http://www.ncsl.org/research/about-state-legislatures/partisan-composition.aspx
# leg <- read_csv("data/legiscontrol_2002_2014.csv")
# leg %<>% 
#   gather(key = "year", value = "state_party", -State) %>%
#   mutate(year = as.integer(year)) %>%
#   mutate(state_party = state_party %>% 
#            str_replace("Split", "Divided") %>% 
#            str_replace("Dem", "Unified Democrat") %>% 
#            str_replace("Rep", "Unified Republican")) %>% 
#   left_join(tibble(State = state.name,
#                        state = state.abb) )
# 
# d %<>% left_join(leg)

# #https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/BSLEFD
# library(haven)
# shor <- read_dta("data/shor+mccarty+1993-2016+state+aggregate+data+May+2018+release+(Updated+July+2018).dta")

#http://ippsr.msu.edu/public-policy/correlates-state-policy
# http://ippsr.msu.edu/sites/default/files/CorrelatesCodebook.pdf
csp <- read.csv("http://ippsr.msu.edu/sites/default/files/correlatesofstatepolicyprojectv2_1.csv") %>% 
  mutate(state = as.character(st))

# ranney4_control = 0 = unified Republican control, 1 = unified Democratic control, .5=neither.
# names(csp)
csp %<>% 
  select(state, year, 
         ranney4_control, 
         govname1, govparty_a, #govparty_b,
         ranney1_sen_dem_prop, ranney1_sen_dem_prop,
         sen_dem_in_sess,
         sen_rep_in_sess,
         sen_ind_in_sess,
         sen_tot_in_sess,
         hs_dem_in_sess,
         hs_rep_in_sess,
         hs_ind_in_sess,
         hs_tot_in_sess) %>%
  distinct()  %>% 
  mutate(state_party = ifelse(ranney4_control == 0, 
                              "All Republican", 
                              ifelse(ranney4_control == 1,
                                     "All Democratic",
                                     "Divided"))) %>% 
  mutate(Gov_Party = ifelse(govparty_a == 0, 
                              "R", 
                              ifelse(govparty_a == 1,
                                     "D",
                                     "Other"))) %>%
  mutate(Leg_Composition = ifelse(hs_rep_in_sess > hs_tot_in_sess/2 & sen_rep_in_sess > sen_tot_in_sess/2, "R", "S")) %>% 
  mutate(Leg_Composition = ifelse(hs_dem_in_sess > hs_tot_in_sess/2 & sen_dem_in_sess > sen_tot_in_sess/2, "D",Leg_Composition )) 

csp %<>% 
  rename(Gov_Name = govname1)
csp %<>% select(state, #statenum, 
               year, Leg_Composition, Gov_Name, Gov_Party, #Gov_Party2,
               state_party) %>% 
  distinct() %>% 
  filter(year > 2000, year < 2017, state != "DC") %>% 
  arrange(state, -year)

#IF WE WANT TO GO WITH LEG CODING THAT CALLS YEARS WHERE POWER TRANSITIONED "SPLIT" (E.G. MS in 2007)
csp %<>%
  mutate(state_party = ifelse(!is.na(Leg_Composition) & Leg_Composition == "R" & Gov_Party == "R", "All Republican", state_party)) %>%
  mutate(state_party = ifelse(!is.na(Leg_Composition) & Leg_Composition == "D" & Gov_Party == "D", "All Democrat", state_party)) %>%
  mutate(state_party = ifelse(!is.na(Leg_Composition) & Leg_Composition != Gov_Party, "Divided", state_party))

csp %<>%
  mutate(Leg_Composition = ifelse(state_party == "All Democratic", "D", Leg_Composition)) %>% 
  mutate(Leg_Composition = ifelse(state_party == "All Republican", "R", Leg_Composition)) %>% 
  mutate(Gov_Party = ifelse(state_party == "All Democratic", "D", Gov_Party)) %>% 
  mutate(Gov_Party = ifelse(state_party == "All Republican", "R", Gov_Party))

write_csv(csp, path = "data/divided_government.csv")

d %<>% left_join(csp)
write_csv(d, path = "data/DIME+divided_government.csv")

```

[Download Divided Government Data](https://raw.githubusercontent.com/judgelord/state-level-DIME/master/data/divided_government.csv)

[Download DIME 2003 and 2007 sample merged with Divided Government Data](https://raw.githubusercontent.com/judgelord/state-level-DIME/master/data/DIME+divided_government.csv)

Source: Correlates of State Policy, [IPPSR.msu.edu](http://ippsr.msu.edu/public-policy/correlates-state-policy)

[Codebook](http://ippsr.msu.edu/sites/default/files/CorrelatesCodebook.pdf)

Variables: state, year, 
         ranney4_control, ranney1_sen_dem_prop, ranney1_sen_dem_prop,
         govname1, govparty_a, govparty_b,
         sen_dem_in_sess,
         sen_rep_in_sess,
         sen_ind_in_sess,
         sen_tot_in_sess,
         hs_dem_in_sess,
         hs_rep_in_sess,
         hs_ind_in_sess,
         hs_tot_in_sess


# By DIME ideology score:

Which state governments are all conservative? Which are all liberal? Which are mixed? 
```{r DIME_plot_function}
## Define a function for plotting state-level DIME data with ggplot
DIMEplot <- function(d){ 
  d %>%
    # define y as state for all layers
  ggplot(aes(y = reorder(state, distance_1_gov_leg_med))) +
    # x = lines for the distance between max and min actors
  geom_segment(aes(x = min, 
                   xend = max, 
                   yend = reorder(state, distance_1_gov_leg_med)), 
               color = "grey", 
               size = 2.1, 
               lineend = "round", 
               alpha = .3) + 
    # x = points for each actor
  geom_point(aes(x = Governor, color = Governor), shape = "G") + 
  geom_point(aes(x = Lower, color = Lower), shape = "L") + 
  geom_point(aes(x = Upper, color = Upper), shape = "U") + 
    # labels 
  labs(y = "States by Preference Divergence", 
       x = "Mean Ideology Score", 
       title = paste("Preference Divergence of Political Principals,", unique(d$year) ), color = "") +
    # scales
  scale_x_continuous(breaks = c(-1, 0, 1)) + 
  scale_colour_gradient2(low = "blue", mid = "black", high = "red") +
    # facet by liberal, conservative, or mixed
  facet_grid(state_ideology ~ ., scales = "free_y", space = "free_y") + 
    # adjust text size and drop legend
  theme(axis.text.y = element_text(size = 5), 
        title = element_text(size = 7), 
        legend.position = "none") 
}
```   

Alignment among governors (G) and upper (U) and lower (L) chambers of state legislatures in 2003 and 2007:
```{r DIME-2003-2007-plots, fig.height = 5, fig.width = 3.1, fig.show='hold'}
d %>% 
  filter(year == 2003) %>% 
  DIMEplot()
d %>% 
  filter(year == 2007) %>%
  DIMEplot()
```     

# By party control and ideology score:

Which state governments are all democrat? Which are all republican? Which are mixed? 
```{r DIME-party_plot_function}
## Define a function for plotting state-level DIME data with ggplot
DIMEplot <- function(d){ 
  d %>%
    # define y as state for all layers
  ggplot(aes(y = reorder(state, distance_1_gov_leg_med))) +
    # x = lines for the distance between max and min actors
  geom_segment(aes(x = min, 
                   xend = max, 
                   yend = reorder(state, distance_1_gov_leg_med)), 
               color = "grey", 
               size = 2.1, 
               lineend = "round", 
               alpha = .3) + 
    # x = points for each actor
  geom_point(aes(x = Governor, color = Governor), shape = "G") + 
  geom_point(aes(x = Lower, color = Lower), shape = "L") + 
  geom_point(aes(x = Upper, color = Upper), shape = "U") + 
    # labels 
  labs(y = "States by Preference Divergence", 
       x = "Mean Ideology Score", 
       title = paste("Preference Divergence of Political Principals,", unique(d$year) ), color = "") +
    # scales
  scale_x_continuous(breaks = c(-1, 0, 1)) + 
  scale_colour_gradient2(low = "blue", mid = "black", high = "red") +
    # facet by liberal, conservative, or mixed
  facet_grid(state_party ~ ., scales = "free_y", space = "free_y") + 
    # adjust text size and drop legend
  theme(axis.text.y = element_text(size = 5), 
        title = element_text(size = 7), 
        legend.position = "none") 
}
```   

Alignment among governors (G) and upper (U) and lower (L) chambers of state legislatures in 2003 and 2007:
```{r DIME-party-2003-2007-plots, fig.height = 6, fig.width = 3.1, fig.show='hold'}
d %>% 
  filter(year == 2003) %>% 
  DIMEplot()
d %>% 
  filter(year == 2007) %>%
  DIMEplot()
```     

